---
title: "Practium_II"
author: "Richard Bird"
date: "8/12/2019"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries
```{r LoadLibraries}
library("R.utils")
library("caret")
library("rpart")
library("rpart.plot")
library("mlbench")
library("corrplot")
library("class")
library("e1071")
library("gmodels")
library("mice")
library("neuralnet")
library("plyr") 
library("pROC")
```

## Multicore
```{r Multicore}
#library("doMC")
#registerDoMC(cores = 4)
```

## Read delimited file to dataframe
```{r ReadDataset}
FileName <- "statistics_export.txt"
BasketballBKP <- read.csv(FileName, head=TRUE, sep=",")
Basketball <- BasketballBKP
```

## Summarize basketball
```{r BasketballSummary}
summary(Basketball)
```

## Summarize dataset
```{r BasketballStr}
str(Basketball)
```

## Count dataset
Number of Rows Basketball
```{r BasketballCount}
nrow(Basketball)
```

## Check dataset for null variable entries and exclude records from dataset with null variable entries.
```{r BasketballMissingValues}
anyNA(Basketball)
```

#
#
#
### DECISION TREE NORMALIZE ###
#
#
#

## Define Normalize function
```{r NormalizeFunction}
normalize<-function(x) {return (((x - min(x)) / (max(x) - min(x))))}
```

```{r BasketballNormalize}
BasketballNormalize <- as.data.frame(lapply(BasketballBKP[1:966], normalize))
anyNA(BasketballNormalize)
BasketballNormalize[is.na(BasketballNormalize)] <- 0
anyNA(BasketballNormalize)
str(BasketballNormalize)
nrow(BasketballNormalize)
```

## Convert label to factor
```{r ConvertLabelToFactorNormalize}
BasketballNormalize$RESULT <- factor(BasketballNormalize$RESULT)
str(BasketballNormalize)
```

#
#
#
### IMPORTANCE DATAFRAME ###
#
#
#

```{R ImportanceDataframe}
BasketballIMP <- cbind(as.data.frame(BasketballNormalize$WON_LOST_PER_VALUE),
                       as.data.frame(BasketballNormalize$WON_LOST_PER_RANK),
                       as.data.frame(BasketballNormalize$OPP_WON_LOST_PER_VALUE),
                       as.data.frame(BasketballNormalize$OPP_WON_LOST_PER_RANK),
                       as.data.frame(BasketballNormalize$SCORING_MARGIN_VALUE),
                       as.data.frame(BasketballNormalize$OPP_SCORING_MARGIN_VALUE),
                       as.data.frame(BasketballNormalize$SCORING_MARGIN_RANK),
                       as.data.frame(BasketballNormalize$HOME_OR_AWAY),
                       as.data.frame(BasketballNormalize$FIELD_GOAL_PER_VALUE),
                       as.data.frame(BasketballNormalize$FIELD_GOAL_PER_RANK),
                       as.data.frame(BasketballNormalize$OPP_SCORING_MARGIN_RANK),
                       as.data.frame(BasketballNormalize$THREE_PT_FIELD_GOALS_PER_RANK),
                       as.data.frame(BasketballNormalize$SCORING_OFFENSE_RANK),
                       as.data.frame(BasketballNormalize$FIELD_GOAL_PER_DEFENSE_RANK),
                       as.data.frame(BasketballNormalize$OPP_GAMES_PLAYED_05),
                       as.data.frame(BasketballNormalize$THREE_PT_FIELD_GOALS_PER_VALUE),
                       as.data.frame(BasketballNormalize$RESULT))
colnames(BasketballIMP) <- c("WON_LOST_PER_VALUE", 
                             "WON_LOST_PER_RANK", 
                             "OPP_WON_LOST_PER_VALUE", 
                             "OPP_WON_LOST_PER_RANK",
                             "SCORING_MARGIN_VALUE",
                             "OPP_SCORING_MARGIN_VALUE",
                             "SCORING_MARGIN_RANK",
                             "HOME_OR_AWAY",
                             "FIELD_GOAL_PER_VALUE",                         
                             "FIELD_GOAL_PER_RANK",
                             "OPP_SCORING_MARGIN_RANK",
                             "THREE_PT_FIELD_GOALS_PER_RANK",
                             "SCORING_OFFENSE_RANK",
                             "FIELD_GOAL_PER_DEFENSE_RANK",
                             "OPP_GAMES_PLAYED_05",
                             "THREE_PT_FIELD_GOALS_PER_VALUE",
                             "RESULT")
str(BasketballIMP)
nrow(BasketballIMP)
anyNA(BasketballIMP)
head(BasketballIMP$RESULT)
```

```{r BasketballTrainTestIMP}
set.seed(4567)
BasketballTrainIndexIMP <- createDataPartition(y = BasketballIMP$RESULT, p = 0.1, list=FALSE)
BasketballTrainIMP <- BasketballIMP[BasketballTrainIndexIMP,]
BasketballTestIMP <- BasketballIMP[-BasketballTrainIndexIMP,]
dim(BasketballTrainIMP)
dim(BasketballTestIMP)
```

#
#
#
### RANDOM FOREST ###
#
#
#

## Basketball Random Forest
```{r BasketballRandomForest}
set.seed(4567)
system.time(BasketballRandomForest <- train(RESULT~., data=BasketballTrainIMP, method="rf"))
```

## Basketball Random Forest Accuracy Print
```{r BasketballRandomForestAccuracy}
BasketballRandomForest$results
print(BasketballRandomForest)
plot(BasketballRandomForest)
```

## Basketball Random Forest
```{r BasketballRandomForestPredict}
system.time(BasketballRandomForestPredict <- predict(BasketballRandomForest, BasketballTestIMP))
```

## Basketball Random Forest confusion matrix.
```{r BasketballRandomForestFinalPredicttTest}
confusionMatrix(BasketballRandomForestPredict, BasketballTestIMP$RESULT)
```

## References

## Lantz, B. (2015). Machine Learning with R. Birmingham, UK: Packt.

## Saxena, R. (2017). Decision Tree Classifier Implementation in R, Retrieved 12:00, June 2, 2019, from http://dataaspirant.com/2017/02/03/decision-tree-classifier-implementation-in-r/ 

## Brownlee, J., R. (2014). Feature Selection with the Caret R Package, Retrieved 12:30, June 2, 2019, from https://machinelearningmastery.com/feature-selection-with-the-caret-r-package/

## Brownlee, J., R. (2016). Tune Machine Learning Algorithms in R (Random Forest Case Study), Retrieved 13:00, June 2, 2019, from https://machinelearningmastery.com/tune-machine-learning-algorithms-in-r/

## Big Computing. (2018). Big Computing, Retrieved 13:30, June 2, 2019, from http://bigcomputing.blogspot.com/2014/10/an-example-of-using-random-forest-in.html

## Guru99. (2018). R Random Forest Tutorial with Example, Retrieved 14:00, June 2, 2019, from https://www.guru99.com/r-random-forest-tutorial.html

## Tahsildar, S. (2019). Gini Index for Decision Trees, Retrieved 14:30, June 2, 2019, from https://blog.quantinsti.com/gini-index/

## RDocumentation. (2019). svm, Retrieved 20:00, June 3, 2019, from https://www.rdocumentation.org/packages/e1071/versions/1.7-1/topics/svm

## RDocumentation. (2019). neuralnet, Retrieved 20:00, June 3, 2019, from https://www.rdocumentation.org/packages/neuralnet/versions/1.44.2/topics/neuralnet

